---
author: Your Name
title: 'Project: Scraping Public Data Screencast'
output: html
---

# Project Description

For your class project (which will take the place of the midterm exam), you will be recording a screencast in the style of David Robinson's TidyTuesday screencasts.

You can find time-stamped, catalogued versions of some of David Robinson's screencasts [here](https://www.rscreencasts.com/). 

Requirements:

- Your screencast should be approximately 45 minutes long.
- Your screencast should begin with identifying and explaining how to scrape or assemble from API calls some public data. This data can be generated by local, state, or national governments, nonprofit organizations, or advocacy organizations, but should not be from any for-profit entity.
- You should showcase at least 4 different techniques you've learned in Stat 351. Some examples include:

    - web scraping or API use
    - efficient/"polite" techniques for acquiring data
    - functional programming
    - list processing
    - use of appropriate and well-chosen graphics
    - interactive graphics
    
Unlike David Robinson's screencasts, you will write a rough pseudocode "script" before you start recording. 
This will give you a rough outline of how to do the analysis and what things you intend to cover.

Your goal is to help a future Stat 351 student understand some of the topics covered in this class. 
So while David Robinson and others who record their screencasts live might not fully explain what he's doing, you should take the time to explain each technique you decide to use in a way that will help someone else understand.


There will be three deliverables for this project:

1. [Plan your dataset and topics](Dataset-Topics.qmd)
2. [Pseudocode script](pseudocode.qmd) uploaded to github repository
3. Screencast + github repository
    - Screencast uploaded to YouTube/YuJa
    - Approximate time index provided for each of the 4 techniques you're demonstrating ([examples](https://www.rscreencasts.com/))
    - Code uploaded to github repository

In lieu of a midterm exam, you will peer review a selection of your classmates' screencasts. 


1. Webscraping 
2. Ethical API usage
3. EDA/LIST PROCESSING
4. Good use of graphics

As I start this the USCF is migrating the site to https://ratings.uschess.org/player/16088033
```{r}
# 1 Setup: Libraries, URLs, and Paths
library(rvest)
library(httr)
library(tidyverse)
library(purrr)
library(tigris)
library(maps)
library(viridisLite)

# Base URLs
SAFE_BASE <- "https://new.uschess.org/safesport-certified-tds?page="
MSA_BASE <- "https://www.uschess.org/msa/MbrDtlMain.php?"

# Cache files
td_cache_file <- "td_list_all.csv"
ratings_cache_file <- "td_ratings_all.csv"

# Ethical delay (1–2 seconds random)
ethical_delay <- function() Sys.sleep(runif(1, 1, 2))
```

```{r}
# 2 Helper: Scrape one SafeSport TD page
scrape_td_page <- function(page_num = 0) {
  url <- paste0(SAFE_BASE, page_num)
  message("Fetching SafeSport page ", page_num, ": ", url)
  
  page <- try(read_html(url), silent = TRUE)
  if (inherits(page, "try-error")) {
    message("Failed to load page ", page_num)
    return(NULL)
  }
  
  tbl <- page |>
    html_element("table") |>
    html_table(fill = TRUE)
  
  if (is.null(tbl) || nrow(tbl) == 0) return(NULL)
  
  names(tbl) <- c("Name", "MemberID", "Expiration", "State", "ActivityType")
  
  tbl |>
    filter(str_detect(MemberID, "^[0-9]+$")) |>
    mutate(MemberID = as.character(MemberID))
}


# 3 Scrape all SafeSport TD page
if (!file.exists(td_cache_file)) {
  all_tds <- list()
  
  for (i in 0:30) { # go through pages until blank
    ethical_delay()
    page_data <- scrape_td_page(i)
    if (is.null(page_data) || nrow(page_data) == 0) break
    all_tds[[length(all_tds) + 1]] <- page_data
  }
  
  td_table <- bind_rows(all_tds)
  write_csv(td_table, td_cache_file)
  message("Saved SafeSport TD list (", nrow(td_table), " rows).")
} else {
  td_table <- read_csv(td_cache_file, show_col_types = FALSE)
  message("Loaded cached TD list (", nrow(td_table), " rows).")
}

# 4 Helper: Scrape MSA rating details for each TD
parse_td_profile <- function(memid) {
  ethical_delay()
  url <- paste0(MSA_BASE, memid)
  
  message("Scraping MSA profile: ", url)
  
  page <- try(read_html(url), silent = TRUE)
  if (inherits(page, "try-error")) {
    message("Error reading ", memid)
    return(tibble(MemberID = memid, State = NA, Regular = NA, Quick = NA, Blitz = NA))
  }
  
  text <- html_text(page)
  
  # Extract State (2-letter code near "State:")
  state <- str_match(text, "(?i)State[:\\s]+([A-Z]{2})")[,2]
  
  # Extract Ratings (more robust regex)
  reg <- str_match(text, "(?i)Regular[\\s\\n]*Rating[:\\s]*([0-9]{3,4})")[,2]
  quick <- str_match(text, "(?i)Quick[\\s\\n]*Rating[:\\s]*([0-9]{3,4})")[,2]
  blitz <- str_match(text, "(?i)Blitz[\\s\\n]*Rating[:\\s]*([0-9]{3,4})")[,2]
  
  tibble(
    MemberID = memid,
    State = state,
    Regular = as.numeric(reg),
    Quick = as.numeric(quick),
    Blitz = as.numeric(blitz)
  )
}

# 5 Scrape all MSA profiles with caching
if (!file.exists(ratings_cache_file)) {
  message("Scraping all TD rating profiles...")
  ratings_data <- map_dfr(td_table$MemberID, parse_td_profile)
  write_csv(ratings_data, ratings_cache_file)
  message("Ratings data saved (", nrow(ratings_data), " rows).")
} else {
  existing <- read_csv(ratings_cache_file, show_col_types = FALSE)
  to_scrape <- setdiff(td_table$MemberID, existing$MemberID)
  
  if (length(to_scrape) > 0) {
    message("Updating ratings for new TDs (", length(to_scrape), " new)...")
    new_data <- map_dfr(to_scrape, parse_td_profile)
    ratings_data <- bind_rows(existing, new_data)
    write_csv(ratings_data, ratings_cache_file)
  } else {
    ratings_data <- existing
    message("Ratings cache up to date (", nrow(ratings_data), " total).")
  }
}

# 6 Merge SafeSport + MSA data
td_data <- left_join(td_table, ratings_data, by = "MemberID") |>
  filter(!is.na(Regular) | !is.na(Quick) | !is.na(Blitz))
```

```{r}
td_data <- td_data %>%
  select(Name, MemberID, State = State.x, Regular, Quick, Blitz)
```

```{r}
# Drop rows missing all ratings or with unrealistic values
td_data <- td_data %>%
  filter(!(is.na(Regular) & is.na(Quick) & is.na(Blitz))) %>%
  mutate(across(c(Regular, Quick, Blitz),
                ~ ifelse(. < 100 | . > 3000, NA, .)))

td_data <- td_data |>
  distinct(MemberID, .keep_all = TRUE)

# ==========================================
# 7 Compute State Averages
# ==========================================
state_avg <- td_data %>%
  group_by(State) %>%
  summarise(
    Regular = mean(Regular, na.rm = TRUE),
    Quick = mean(Quick, na.rm = TRUE),
    Blitz = mean(Blitz, na.rm = TRUE),
    Count = n()
  ) %>%
  filter(!is.na(State) & State %in% state.abb)

# ==========================================
# 8 Visualizations
# ==========================================
us_states <- map_data("state")
state_names <- data.frame(State = state.abb, region = tolower(state.name))

merged <- left_join(state_avg, state_names, by = "State") %>%
  left_join(us_states, by = "region")

# (a) U.S. Map of Average Regular Rating
p_map <- ggplot(merged, aes(long, lat, group = group, fill = Regular, alpha = sqrt(Count))) +
  geom_polygon(color = "grey40", size = 0.2) +
  scale_fill_viridis_c(option = "plasma", direction = -1, na.value = "grey90") +
  scale_alpha_continuous(range = c(0.5, 1)) +
  coord_fixed(1.3) +
  theme_void() +
  labs(
    title = "Average Regular TD Rating by State (SafeSport-Certified TDs)",
    subtitle = "Opacity reflects number of certified TDs per state",
    fill = "Avg Rating",
    alpha = "TD Count (sqrt-scaled)",
    caption = paste("Data scraped:", Sys.Date())
  )

ggsave("us_td_rating_map.png", plot = p_map, width = 10, height = 6)

p_map
```

```{r}
# (b) Bar Chart: Time Control Comparison
avg_ratings <- colMeans(state_avg[, c("Regular", "Quick", "Blitz")], na.rm = TRUE)
df_avg <- tibble(TimeControl = names(avg_ratings), Rating = avg_ratings)

p_bar <- ggplot(df_avg, aes(reorder(TimeControl, Rating), Rating, fill = TimeControl)) +
  geom_col(width = 0.6, show.legend = FALSE) +
  geom_text(aes(label = round(Rating, 1)), vjust = -0.5, fontface = "bold") +
  scale_fill_viridis_d(option = "turbo") +
  theme_minimal(base_size = 13) +
  labs(
    title = "Average Ratings by Time Control",
    x = NULL, y = "Average Rating"
  ) +
  ylim(0, max(df_avg$Rating) * 1.1)

p_bar
```

```{r}
# (c) TD Count per State
top_states <- state_avg %>%
  arrange(desc(Count)) %>%
  slice_head(n = 20)

p_count <- ggplot(top_states, aes(x = reorder(State, Count), y = Count)) +
  geom_col(fill = viridis(20)[10]) +
  geom_text(aes(label = Count), hjust = -0.1, size = 3.5) +
  coord_flip() +
  theme_minimal(base_size = 12) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  labs(
    title = "Top 20 States by Number of Certified TDs",
    x = "State", y = "TD Count"
  )
p_count
ggsave("td_count_by_state.png", plot = p_count, width = 7, height = 6)
```
```{r}
top_by_state <- function(state_abbrev, data = td_data) {
  
  # Basic validation
  if (!state_abbrev %in% unique(data$State)) {
    stop("State abbreviation not found in dataset: ", state_abbrev)
  }
  
  # Filter, sort, take top 10
  top10 <- data |>
    filter(State == state_abbrev) |>
    arrange(desc(Regular)) |>
    slice_head(n = 10)
  
  return(top10)
}

top_by_state("CT")
```
```{r}
top10_nationwide <- td_data |>
  filter(!is.na(Regular)) |>
  arrange(desc(Regular)) |>
  slice_head(n = 10)

top10_nationwide
```

```{r}
# 9 Summary Report
cat("Total unique TDs analyzed: ", nrow(td_data), "\n")
cat("States represented:        ", n_distinct(td_data$State), "\n")

overall_stats <- td_data %>%
  summarise(
    Avg_Regular = mean(Regular, na.rm = TRUE),
    Med_Regular = median(Regular, na.rm = TRUE),
    SD_Regular = sd(Regular, na.rm = TRUE),
    Avg_Quick = mean(Quick, na.rm = TRUE),
    Med_Quick = median(Quick, na.rm = TRUE),
    SD_Quick = sd(Quick, na.rm = TRUE),
    Avg_Blitz = mean(Blitz, na.rm = TRUE),
    Med_Blitz = median(Blitz, na.rm = TRUE),
    SD_Blitz = sd(Blitz, na.rm = TRUE)
  )

print(overall_stats)
```

```{r}
top5 <- state_avg %>% arrange(desc(Regular)) %>% slice_head(n = 5)
print(top5)
```

```{r}
bottom5 <- state_avg %>% arrange(Regular) %>% slice_head(n = 5)
print(bottom5)
```

```{r}
largest <- state_avg %>% arrange(desc(Count)) %>% slice_head(n = 5)
print(largest)
```

